ICM testing

### 2023-06-02 Scripts/ASR_stocks.Rmd
Many many updates, multi-species NS model has been developed
The NS multi-species model is in its infancy, but if we can land a student it can certainly be generalized nicely.

  N.end <- sum(data[data$Year==max(years),]$Num)
  #N.end <- rowSums(abund[nrow(abund),-which(names(abund) %in% c("Year", "Stock", "type"))], na.rm=T)
  #vpa.abund <- rowSums(abund[,-which(names(abund) %in% c("Year", "Stock", "type"))], na.rm=T)
  vpa.abund <- data %>% dplyr::group_by(Year) %>% dplyr::summarize(vpa=sum(Num)) %>% pull(vpa)

  # The real mx matrix, recruits produced per individual in each age class... Not perfect as I need to offset recruits/ssb, but close enough for the moment..
  #minage
  minage <- min(as.numeric(data$age))
  #maxage
  maxage <- max(as.numeric(data$age))
  #recruits
  annual <- data.frame(Year=data$Year[data$age==minage], recruits=data$Num[data$age==minage])
  #ssn
  data$ssn <- data$Num * data$AM
  #ssb
  data$ssb <- data$ssn * data$WA #I need to figure out how to line this up with the number of recruits for the right years in the data object.... got it right for the overall r.p.ssb below.
  #tot.ssb
  annual <- data %>% group_by(Year) %>% summarize(tot.ssb = sum(ssb)) %>% left_join(annual)
  #r.p.ssb This needs to be offset by the age of recruits
  annual$r.p.ssb <- c(rep(NA,minage),annual$recruits[(minage+1):nrow(annual)]/annual$tot.ssb[1:(nrow(annual)-minage)])
  # recs.per.age
  data <- left_join(data, annual)
  # So I need to make an offset SSB in the 'data' object to line up with the correct r.p.ssb field.  This is probably gonna suck...
  tmp <- NULL
  for(j in 1:length(years))
  {
    tst <- data %>% dplyr::filter(Year == years[j]) %>% dplyr::select(ssb,ssn,Year,age)
    tst$Year <- tst$Year + minage
    names(tst) <- c("ssb.offset","ssn.offset","Year","age")
    tmp[[j]] <- tst
  }
  # Unpack the list
  ssb.off <- do.call('rbind',tmp)
  # And merge it with the data object
  data <- left_join(data,ssb.off,by=c("Year","age"))
  data$recs.per.age <- data$ssb.offset*data$r.p.ssb
  # mx
  # Some of the stocks are just the males and females, so these are already half the population.  Probably can just not include the male stocks in the end.
  # This works because 'males' is also in 'females'
  if(grepl("males",i)) data$mx <- data$recs.per.age/data$ssn.offset 
  if(!grepl("males",i))data$mx <- data$recs.per.age/data$ssn.offset/2 # Moms only! 
  data$mx[is.nan(data$mx)] <- 0 # if we don't have any spawners in an age class in a year their fecundity is 0
  # Easier to remove the years where we don't have recs.per.age...
  #data <- data %>% dplyr::filter(Year %in% years[(minage+1):length(years)])
  # Something is wrong with the decline rate method, but the exponential and logistic are working pretty... pretty... pretty good...

@@

back.res <- NULL
back.ts <- NULL
back.ts.quants <- NULL
tuned.res <- NULL
for(i in Stocks)
{
  years <- years.tmp[[i]]
  prop.nat.mort <- pnm.tmp[[i]] 
  weight.age <- waa.tmp[[i]] 
  ages <- ages.tmp[[i]] 
  rem <- rem.tmp[[i]] 
  mx <- mx.tmp[[i]] 
  N.end <- NE.tmp[[i]] 
  vpa.abund <- vpa.tmp[[i]] 
  age.mat <- am.tmp[[i]]

  tst <- icm.sim(years,
                 mat.age = age.mat,
                 nm = prop.nat.mort,
                 w.age = weight.age,
                 ages = ages,
                 rems = rem$rem,
                 fecund = mx,
                 N.end = N.end,
                 pop.model = 'exponential',
                 n.sims = 500,
                 sd.mat = 0.5,
                 sd.nm = 0.5,
                 sd.wt = 0.5,
                 sd.fecund = 0.5)

@@

# Now run the tuning sims

  N.end <- vpa.abund[length(vpa.abund)]
  tunes <- tune.sim(years,
                    n.steps = 100,
                    abund.ts = vpa.abund,
                    mat.age = age.mat,
                    nm = prop.nat.mort,
                    w.age = weight.age,
                    ages = ages,
                    rems = rem$rem,
                    fecund = mx,
                    N.end = N.end,
                    pop.model = 'exponential')

  tuned.res[[i]] <- tunes
} # end i in Stocks look
# Save all these
saveRDS(object = tuned.res,file = "D:/Github/ICM/Results/tuned_fits.Rds")
saveRDS(object = back.res,file = "D:/Github/ICM/Results/backwards_sim_results.Rds")
saveRDS(object = back.ts,file = "D:/Github/ICM/Results/backwards_sim_ts_summaries.Rds")
saveRDS(object = back.ts.quants,file = "D:/Github/ICM/Results/backwards_sim_ts_quantiles_summaries.Rds")
```

@@

```{r, forward-sims-n-figs,echo=F, message=F, warning=F, fig.width=6, fig.height=4}
```

>>>>

  N.end <- sum(data[data$Year==max(years),]$Num)
  vpa.abund <- data %>% dplyr::group_by(Year) %>% dplyr::summarize(vpa=sum(Num,na.rm=T)) %>% pull(vpa)

  # The real mx matrix, recruits produced per individual in each age class... Not perfect as I need to offset recruits/ssb, but close enough for the moment..
  #minage
  minage <- min(as.numeric(data$age))
  #maxage
  maxage <- max(as.numeric(data$age))
  #recruits
  annual <- data.frame(Year=data$Year[data$age==minage], recruits=data$Num[data$age==minage])
  #ssn
  data$ssn <- data$Num * data$AM
  #ssb
  data$ssb <- data$ssn * data$WA #I need to figure out how to line this up with the number of recruits for the right years in the data object.... got it right for the overall r.p.ssb below.
  #tot.ssb
  annual <- data %>% group_by(Year) %>% summarize(tot.ssb = sum(ssb)) %>% left_join(annual)

  #tst <- data %>% group_by(Year) %>% summarize(tot.ssn = sum(ssn)) %>% left_join(annual)

  #r.p.ssb This needs to be offset by the age of recruits
  annual$r.p.ssb <- c(rep(NA,minage),annual$recruits[(minage+1):nrow(annual)]/annual$tot.ssb[1:(nrow(annual)-minage)])
  # recs.per.age
  data <- left_join(data, annual)
  # So I need to make an offset SSB in the 'data' object to line up with the correct r.p.ssb field.  This is probably gonna suck...
  tmp <- NULL
  for(j in 1:length(years))
  {
    tst <- data %>% dplyr::filter(Year == years[j]) %>% dplyr::select(ssb,ssn,Year,age)
    tst$Year <- tst$Year + minage
    names(tst) <- c("ssb.offset","ssn.offset","Year","age")
    tmp[[j]] <- tst
  }
  # Unpack the list
  ssb.off <- do.call('rbind',tmp)
  # And merge it with the data object
  data <- left_join(data,ssb.off,by=c("Year","age"))
  data$recs.per.age <- data$ssb.offset*data$r.p.ssb
  # mx
  # Some of the stocks are just the males and females, so these are already half the population.  Probably can just not include the male stocks in the end.
  # This works because 'males' is also in 'females'
  if(grepl("males",i)) data$mx <- data$recs.per.age/data$ssn.offset 
  if(!grepl("males",i))data$mx <- data$recs.per.age/data$ssn.offset/2 # Moms only! 
  data$mx[is.nan(data$mx)] <- 0 # if we don't have any spawners in an age class in a year their fecundity is 0
  # Easier to remove the years where we don't have recs.per.age...
  #data <- data %>% dplyr::filter(Year %in% years[(minage+1):length(years)])
  # Something is wrong with the decline rate method, but the exponential and logistic are working pretty... pretty... pretty good...

@@

back.res <- NULL
back.ts <- NULL
back.ts.quants <- NULL
for.res <- NULL
for.ts <- NULL
for.ts.quants <- NULL
tuned.res <- NULL
for(i in Stocks)
{
  years <- years.tmp[[i]]
  prop.nat.mort <- pnm.tmp[[i]] 
  weight.age <- waa.tmp[[i]] 
  ages <- ages.tmp[[i]] 
  rem <- rem.tmp[[i]] 
  mx <- mx.tmp[[i]] 
  N.end <- NE.tmp[[i]] 
  vpa.abund <- vpa.tmp[[i]] 
  age.mat <- am.tmp[[i]]

  tst <- back.sim(years,
                 mat.age = age.mat,
                 nm = -(log(1-prop.nat.mort)),
                 w.age = weight.age,
                 ages = ages,
                 rems = rem$rem,
                 fecund = mx,
                 N.end = N.end,
                 pop.model = 'exponential',
                 n.sims = 500,
                 sd.mat = 0.5,
                 sd.nm = 0.5,
                 sd.wt = 0.5,
                 sd.fecund = 0.5)

@@

# ################# Now run as a forward simulation
#   N.start <- vpa.abund[1] 
# 
#   tst <- for.sim(years,
#                  mat.age = age.mat,
#                  nm = -(log(1-prop.nat.mort)),
#                  w.age = weight.age,
#                  ages = ages,
#                  rems = rem$rem, # rep(0,nrow(rem)),
#                  fecund = mx,
#                  N.start = N.start,
#                  pop.model = 'exponential', K = 2e6,
#                  n.sims = 2,
#                  sd.mat = 0.5,
#                  sd.nm = 0.5,
#                  sd.wt = 0.5,
#                  sd.fecund = 0.5)
# 
#   #Combine the data
#   did.it.work <- data.frame(abund = c(vpa.abund,tst$Pop$abund),years = c(years,tst$Pop$years),sim = c(rep('VPA',length(years)),tst$Pop$sim))
#   # Here get the Upper and lower 50% quantiles to make a functional boxplot
#   quants <- did.it.work %>% dplyr::filter(sim != "VPA") %>% dplyr::group_by(years) %>% dplyr::summarise(L.50 = quantile(abund,probs=c(0.25)),
#                                                                         med = median(abund),
#                                                                         U.50 = quantile(abund,probs=c(0.75)))
# 
#   for.res[[i]] <- tst
#   for.ts[[i]] <- did.it.work
#   for.ts.quants[[i]] <- quants

# ########################## Now run the tuning sims
#
   N.end <- vpa.abund[length(vpa.abund)]
   tunes <- tune.sim(years,
                     n.steps = 100,
                     abund.ts = vpa.abund,
                     mat.age = age.mat,
                     nm = -(log(1-prop.nat.mort)),
                     w.age = weight.age,
                     ages = ages,
                     rems = rem$rem,
                     fecund = mx,
                     N.end = N.end,
                     pop.model = 'exponential')

   tuned.res[[i]] <- tunes
} # end i in Stocks look

# Save all these
 saveRDS(object = tuned.res,file = "D:/Github/ICM/Results/tuned_fits.Rds")
 saveRDS(object = back.res,file = "D:/Github/ICM/Results/backwards_sim_results.Rds")
 saveRDS(object = back.ts,file = "D:/Github/ICM/Results/backwards_sim_ts_summaries.Rds")
 saveRDS(object = back.ts.quants,file = "D:/Github/ICM/Results/backwards_sim_ts_quantiles_summaries.Rds")
```

@@

```{r, forward-sims-n-figs,echo=F, message=F, warning=F, fig.width=6, fig.height=4}
# Now we have to forward project the stocks, I think it makes the most sense to do this using the tuned data at the moment
# So we assume the fecundity data and the tuned natural mortality data for the stocks is a reasonable representation of the future
load(file = "D:/Github/ICM/Results/model_inputs.Rdata")

back.res <- NULL
back.ts <- NULL
back.ts.quants <- NULL
for.res <- NULL
for.ts <- NULL
for.ts.quants <- NULL
tuned.res <- NULL
for(i in Stocks)
{
  years <- years.tmp[[i]]
  prop.nat.mort <- pnm.tmp[[i]] 
  weight.age <- waa.tmp[[i]] 
  ages <- ages.tmp[[i]] 
  rem <- rem.tmp[[i]] 
  mx <- mx.tmp[[i]] 
  N.end <- NE.tmp[[i]] 
  vpa.abund <- vpa.tmp[[i]] 
  age.mat <- am.tmp[[i]]

################# Now run as a forward simulation
  N.start <- vpa.abund[length(vpa.abund)] 
  years <- 2025:2075
  # For the forward simulations I'm setting it up so we feed in the mean vector or a character, we'll then create a vector of mx's with variablitity
  st.time <- Sys.time()
  tst <- for.sim(years,
                 mat.age = age.mat,
                 nm = -(log(1-prop.nat.mort)),
                 w.age = weight.age,
                 ages = ages,
                 rems = c(0.1,0.2), # This is set up as a fishing mortality that averages 0.1 with an sd of 0.2)
                 fecund = mx,
                 N.start = N.start,
                 pop.model = 'exponential', sim = 'project', proj.sim = 'sample',
                 n.sims = 30,
                 sd.mat = 0.1,
                 sd.nm = 0.1,
                 sd.wt = 0.1,
                 sd.fecund = 0.1)
  Sys.time() - st.time
  #Combine the data
  did.it.work <- data.frame(abund = tst$Pop$abund,years = c(tst$Pop$years),sim = tst$Pop$sim)
  # Here get the Upper and lower 50% quantiles to make a functional boxplot
  quants <- tst$Pop %>% dplyr::group_by(years) %>% dplyr::summarise(L.50 = quantile(abund,probs=c(0.25)),
                                                                   med = median(abund),
                                                                   U.50 = quantile(abund,probs=c(0.75)))

  for.res[[i]] <- tst
  for.ts[[i]] <- tst$Pop
  for.ts.quants[[i]] <- quants  
}  
ggplot(quants) + geom_line(aes(x=years,y=med))  + geom_ribbon(aes(x=years,ymax=U.50,ymin = L.50),alpha=0.5,fill='blue',color='blue')
